Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-09 paper/paper.tex
Changes 2sum figure to be algorithmic
Experiments

@@

\section{Floating Point}

TODO
Explain 2sum

\begin{figure}
 \begin{lstlisting}[language=c]
s  = a + b;
b2 = s - a;
a2 = s - b2;
db = b - b2;
da = a - a2;
t  = da + db;
 \end{lstlisting}

\begin{lstlisting}[language=c]
s = a + b;
z = s - a;
t = b - z;
\end{lstlisting}

 \caption{2Sum}
  \label{fig:2sum}

\section{Experimental Results}
\label{sec:experiments}

We performed our experiments on a 4-core, 3.07\,GHz Xeon X5667 with 50\,GB of RAM.

TODO
Explain experiments
Explain difference to Brahma (to explain timeouts)
Run experiments on Sketch and AHA

>>>>

\algrenewcommand\algorithmicrequire{\textbf{Precondition:}}
\algrenewcommand\algorithmicensure{\textbf{Postcondition:}}

@@

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc 2Sum
 \label{alg:2sum}}
\begin{algorithmic}
\Let{$s$}{$a+b$}
\Let{$b'$}{$s - a$}
\Let{$a'$}{$s - b'$}
\Let{$\delta b$}{$b - b'$}
\Let{$\delta a$}{$a - a'$}
\Let{$t$}{$\delta a + \delta b$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc Fast2Sum
 \label{alg:fast2sum}}
\begin{algorithmic}
\Require{$a \ge b$}
\Let{$s$}{$a + b$}
\Let{$z$}{$s - a$}
\Let{$t$}{$b - z$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

 \caption{The {\sc 2Sum} algorithm for correctly rounded floating point sums.}
  \label{fig:2sum}
\end{figure}

\section{Experimental Results}
\label{sec:experiments}
We implemented our synthesis procedure as the {\sc kalashnikov} tool.  The implementation consists
of around 600 lines of \newC and 600 lines of Python.  To evaluate the tool we used the 25 bitvector
programs from~\cite{brahma}.  The code we used to perform the experiments, along with the
benchmarks, is available at \texttt{http://www.cprover.org/kalashnikov}.  As a backend solver, we used
{\sc CBMC} at SVN revision r2777, using Glucose 2.2~\cite{glucose} as the SAT solver.
We performed our experiments on a 4-core, 3.07\,GHz Xeon X5667 with 50\,GB of RAM.

To give a reference point for the efficiency of our tool, we present the results given for {\sc brahma}
on the same benchmarks, as reported in~\cite{brahma}.  These experiments were performed an 8-core,
1.86\,GHz Xeon with 4\,GB of RAM, so the timings should be comparable to within a factor of 2 or so.

The results of our experiments are shown in Fig.~\ref{fig:results-table}.  On the problems for which
we are able to synthesise a program, we take a similar time to {\sc brahma}.  We are unable to synthesise
programs for 6 of the 25 problems that {\sc brahma} is able to solve.  We believe this to be a result of
the slight difference in the problems solved by {\sc kalashnikov} and {\sc brahma}.  We do not ask the
user for any information beyond the program specification, whereas {\sc brahma} asks for an extended
component library when synthesis with the standard library fails.  This extended library contains a
reasonable amount of information about the structure of the final program, which we believe makes
problems P19 to P25 tractable for {\sc brahma} but intractable for us.

To demonstrate the capabilities of \newC as an impementation language, we wanted to synthesise a
non-trivial floating point program.  For this we chose the classic {\sc 2Sum} algorithm for computing
exact sums of floating point numbers.  The original {\sc 2Sum} algorithm is shown in Alg.~\ref{alg:2sum}.
In the case that $a \ge b$, exact addition can be implemented in fewer instructions -- this is the
{\sc Fast2Sum} algorithm~\cite{fast2sum}.  We used {\sc 2sum} as our specification, and
added an assumption that $a \ge b$.  {\sc Kalashnikov} was able to synthesise and verify the
code for {\sc Fast2Sum} from this specification in 396.07s, which we believe to be the first time
this has been achieved.  This provides an alternative proof to~\cite{fast2sum} that {\sc Fast2Sum} is
minimal in terms of floating point operations.
